Make any stupid plot you want

### 2021-02-25 draft/sections/implementation.tex
point & line assembly args

\begin{document}
\section{Prototype Implementation: Matplottoy}
\label{sec:implementation}
\begin{figure}[H]
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=\textwidth]{figures/code/scatter_0.png}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=\textwidth]{figures/code/line_1.png}
    \end{subfigure}
    \caption{Scatter plot and line plot implemented using prototype artists and data models, building on Matplotlib rendering. \note{update with bar}}
    \label{fig:code_scatter_line}
\end{figure}

To prototype our model, we implemented the artist classes for the scatter and line plots shown in figure~\ref{fig:code_scatter_line} because they differ in every attribute: different visual channels \vchannel\ that composite to different marks \vmark\ with different continuities \vindex\.  We make use of the Matplotlib figure and axes artists \cite{hunterArchitectureOpenSource,hunterMatplotlib2DGraphics2007} so that we can initially focus on the data to graphic transformations.

@@

We then add the \mintinline{python}{Point} and \mintinline{python}{Line} artist that construct the scatter and line graphics. These artists are implemented as the equivalence class \vartisteq\ with the aesthetic configurations factored out into a \mintinline{python}{transforms} dictionary that specifies the visual bundle \vtotal\. The equivalence classes \vartisteq\ map well to Python classes since the functional aspects-\vchannel, \vmarkd, and \vindex - are completely reusable in a consistent composition, while the visual values in \vtotal\ are what change between different artists belonging to the same class \vartisteq. The \mintinline{python}{data} object is an abstraction of a data bundle \dtotal\ with a specified section \tau. Implementing \gtotal\ and \gsection\ are out of scope for this prototype because they are part of the rendering process. We also did not implement any form of \vindex\ because the scatter, line, and bar plots prototyped here directly broadcast from \dbasepoint to \gbasepoint, unlike for example an image which may need to be rotated.

@@

    def assemble(self, visual):
        # set the properties of the graphic

    def draw(self, renderer):
        # returns K, indexed on fiber then key 
        # is passed the 
        view = self.data.view(self.axes) 
        # visual channel encoding applied fiberwise 
        visual = {p: encoder(view[f] if f is not None else None) 
                     for p, (f, encoder) in self.transforms.items()}
        self.assemble(visual)
        # pass configurations off to the renderer
        super().draw(renderer)
\end{minted}

The data is fetched in section \dsection\ via a \mintinline{python}{view} method on the data because the input to the artist is a section on \dtotal. The \mintinline{python}{view} method takes the \mintinline{python}{axes} attribute because it provides the region in graphic coordinates \gbase\ that we can use to query back into data to select a subset as discussed in section~\ref{sec:data_sheaf_stalk}. The \vchannel\ functions are then applied to the data to generate the visual section \vsection\ that here is the object \mintinline{python}{visual}. We allow for fixed visual parameter, such as setting a constant color for all sections, via setting in \mintinline{python}{None} as the data fiber \dfiber\ name in the \mintinline{python}{transform} dictionary.

The visual object is then passed into the \mintinline{python}{assembly} function that is \vmarkd. This assembly function is responsible for generating a representation of the glyph such that it could be serialized to recreate a static version of the graphic. Although \mintinline{python}{assemble} could be implemented outside the class such that it returns an object the artist could then parse to set attributes, the attributes are directly set here to reduce indirection. This artist is not optimized because we prioritized demonstrating the separability of \vchannel\ and \vmarkd. The last step in the artist function is handing itself off to the renderer. The extra \mintinline{python}{*arg, **kwargs} arguments in \mintinline{python}{__init__,draw} are artifacts of how these objects are currently implemnted in Matplotlib.

The \mintinline{python}{Point} artist builds on \mintinline{python}{collection} artists because collections are optimized to efficiently draw a sequence of primitive point and area marks. In this prototype, the scatter marker shape is fixed as a circle, and the only visual fiber components are x and y position, size, and the facecolor of the marker. We only show the \mintinline{python}{assembly} function here because the \mintinline{python}{__init__, draw} are identical the prototype artist shown in \ref{code:artist}. 
\begin{minted}{python}
class Point(mcollections.Collection):=
    def assemble(self, visual):
        # construct geometries of the circle marks in visual coordinates
        self._paths = [mpath.Path.circle(center=(x,y), radius=s) 
                       for (x, y, s) in zip(visual['x'],visual['y'], visual['s'])] 
        # set attributes of marks, these are vectorized 
        # circles and facecolors are lists of the same size
        self.set_facecolors(visual['facecolors'])

\end{minted} 
The \mintinline{python}{view} method repackages the data as a fiber component indexed table of vertices, as described in section~\ref{sec:triangulization}. Even though the \mintinline{python}{view} is fiber indexed, each vertex at an index \dbasepoint has corresponding values in section $\dsection(\dbasepoint_{i})$. This means that all the data on one vertex maps to one glyph. To ensure the integrity of the section, \mintinline{python}{view} must be atomic. This means that the values cannot change after the method is called in draw until a new call in draw. We put this constraint on the return of the \mintedinline{python}{view} method so that we do not risk race conditions. Using triangulization provides a common interface to the data, one we will reuse for all the artists presented here.

This table is converted to a table of visual variables and is then passed into \mintinline{python}{assemble}. In  \mintinline{python}{assemble}, the \vsection\ is used to individually construct the vector path of each circular marker with center \texttt{(x,y)} and size \texttt{x} and set the colors of each circle. This is done via the \mintinline{python}{Path.circle} object. As mentioned in sections ~\ref{sec:data_sheaf_stalk} and \ref{sec:artist_q}, this assembly function could as easily be implemented such that it was fed one $\dsection(\dbasepoint)$ at a time.

The main difference between the \mintinline{python}{Point} and \mintinline{python}{Line} objects is in th3 \mintinline{python}{assemble} function because line has different continuity from scatter and is represented by a different type of graphical mark.

\begin{minted}{python}
class Line(mcollections.LineCollection):
    def assemble(self, visual):
        #assemble line marks as set of segments 
        segments = [np.vstack((vx, vy)).T for vx, vy 
                    in zip(visual['x'], visual['y'])]
        self.set_segments(segments)
        self.set_color(visual['color'])
\end{minted}
In the \mintinline{python}{Line} artist, \mintinline{python}{view} returns a table of edges. Each edge consists of (x,y) points sampled along the line defined by the edge and information such as the color of the edge. As with \mintinline{python}{Point}, the data is then converted into visual variables. In \mintinline{python}{assemble}, this visual representation is composed into a set of line segments, where each segement is the array generated by \mintinline{python}{np.vstack((vx, vy))}. Then the colors of each line segment are set. The colors are guaranteed to correspond to the correct segment because of the atomicity constraint on view.

\begin{figure}[H]
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=\textwidth]{figures/code/bar_v.png}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=\textwidth]{figures/code/bar_h.png}
    \end{subfigure}
    \caption{Frequency of Penguin types visualized as discrete bars. }
    \label{fig:code_bar_simple}
\end{figure}
The bar charts in figure~\ref{fig:code_bar_simple} are generated with a \mintinline{python}{Bar} artist. The have the same required \vfiber\ components of (position, length). In \mintinline{pyton}{__init__} of \mintinline{python}{Bar} an additional parameter is set, \texttt{orientation} which controls whether the bars are arranged vertically or horizontally and only applies holistically to the graphic and never to individual data parameters. This highlights another advantage of this model, that it encourages explicit differentiation between parameters in \vtotal\ and graphic parameters that specify \vmarkd\.

>>>>

\begin{document}
\section{Prototype Implementation: Matplottoy}
\label{sec:implementation}
\begin{figure}[H]
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=\textwidth]{figures/code/scatter_0.png}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=\textwidth]{figures/code/line_1.png}
    \end{subfigure}
    \caption{Scatter plot and line plot implemented using prototype artists and data models, building on Matplotlib rendering.}
    \label{fig:code_scatter_line}
\end{figure}
To prototype our model, we implemented the artist classes for the scatter and line plots shown in figure~\ref{fig:code_scatter_line} because they differ in every attribute: different visual channels \vchannel\ that composite to different marks \vmark\ with different continuities \vindex\.  We make use of the Matplotlib figure and axes artists \cite{hunterArchitectureOpenSource,hunterMatplotlib2DGraphics2007} so that we can initially focus on the data to graphic transformations.

@@

We then add the \mintinline{python}{Point} and \mintinline{python}{Line} artist that construct the scatter and line graphics. These artists are implemented as the equivalence class \vartisteq\ with the aesthetic configurations factored out into a \mintinline{python}{transforms} dictionary that specifies the visual bundle \vtotal\. The equivalence classes \vartisteq\ map well to Python classes since the functional aspects-\vchannel, \vmarkd, and \vindex - are completely reusable in a consistent composition, while the visual values in \vtotal\ are what change between different artists belonging to the same class \vartisteq. The \mintinline{python}{data} object is an abstraction of a data bundle \dtotal\ with a specified section \dsection. Implementing \gtotal\ and \gsection\ are out of scope for this prototype because they are part of the rendering process. We also did not implement any form of \vindex\ because the scatter, line, and bar plots prototyped here directly broadcast from \dbasepoint\ to \gbasepoint, unlike for example an image which may need to be rotated.

@@

    def assemble(self, **args):
        # set the properties of the graphic

    def draw(self, renderer):
        # returns K, indexed on fiber then key 
        # is passed the 
        view = self.data.view(self.axes) 
        # visual channel encoding applied fiberwise 
        visual = {p: encoder(view[f] if f in self.data.FB.F else f) 
                  for p, (f, encoder) in self.transforms.items()}
        self.assemble(**visual)
        # pass configurations off to the renderer
        super().draw(renderer)
\end{minted}

The data is fetched in section \dsection\ via a \mintinline{python}{view} method on the data because the input to the artist is a section on \dtotal. The \mintinline{python}{view} method takes the \mintinline{python}{axes} attribute because it provides the region in graphic coordinates \gbase\ that we can use to query back into data to select a subset as discussed in section~\ref{sec:data_sheaf_stalk}. The \vchannel\ functions are then applied to the data to generate the visual section \vsection\ that here is the object \mintinline{python}{visual}. We allow for fixed visual parameter, such as setting a constant color for all sections, by verifying that the named component is in \dfiber\ before accessing the data. If the data component name is not in \dfiber\, this is interpreted to mean this component is a thickening of \vtotal that could be pulled back to \dtotal\ via an inverse identity \vchannel. The conversion from data to visual space is simplified here to show that it is the encoding \vchannel\ applied to the component.

The components of the visual object, denoted by the Python unpacking convention \mintinline{python}{**visual} are then passed into the \mintinline{python}{assembly} function that is \vmarkd. This assembly function is responsible for generating a representation of the glyph such that it could be serialized to recreate a static version of the graphic. Although \mintinline{python}{assemble} could be implemented outside the class such that it returns an object the artist could then parse to set attributes, the attributes are directly set here to reduce indirection. This artist is not optimized because we prioritized demonstrating the separability of \vchannel\ and \vmarkd. The last step in the artist function is handing itself off to the renderer. The extra \mintinline{python}{*arg, **kwargs} arguments in \mintinline{python}{__init__,draw} are artifacts of how these objects are currently implemnted in Matplotlib.

The \mintinline{python}{Point} artist builds on \mintinline{python}{collection} artists because collections are optimized to efficiently draw a sequence of primitive point and area marks. In this prototype, the scatter marker shape is fixed as a circle, and the only visual fiber components are x and y position, size, and the facecolor of the marker. We only show the \mintinline{python}{assembly} function here because the \mintinline{python}{__init__, draw} are identical the prototype artist shown in \ref{code:artist}. 
\begin{minted}{python}
class Point(mcollections.Collection):
    def assemble(self, x, y, s, facecolors='C0' ):
        # construct geometries of the circle glyphs in visual coordinates
        self._paths = [mpath.Path.circle(center=(xi,yi), radius=si) 
                    for (xi, yi, si) in zip(x, y, s)] 
        # set attributes of glyphs, these are vectorized 
        # circles and facecolors are lists of the same size
        self.set_facecolors(facecolors)
\end{minted} 
The \mintinline{python}{view} method repackages the data as a fiber component indexed table of vertices, as described in section~\ref{sec:triangulization}. Even though the \mintinline{python}{view} is fiber indexed, each vertex at an index \dbasepoint has corresponding values in section $\dsection(\dbasepoint_{i})$. This means that all the data on one vertex maps to one glyph. To ensure the integrity of the section, \mintinline{python}{view} must be atomic. This means that the values cannot change after the method is called in draw until a new call in draw. We put this constraint on the return of the \mintedinline{python}{view} method so that we do not risk race conditions.

This table is converted to a table of visual variables and is then passed into \mintinline{python}{assemble}. In  \mintinline{python}{assemble}, the \vsection\ components are used to construct the vector path of each circular marker with center \texttt{(x,y)} and size \texttt{x} and set the colors of each circle. This is done via the \mintinline{python}{Path.circle} object. As mentioned in sections ~\ref{sec:data_sheaf_stalk} and \ref{sec:artist_q}, this assembly function could as easily be implemented such that it was fed one $\dsection(\dbasepoint)$ at a time.

The main difference between the \mintinline{python}{Point} and \mintinline{python}{Line} objects is in the \mintinline{python}{assemble} function because line has different continuity from scatter and is represented by a different type of graphical mark.

\begin{minted}{python}
class Line(mcollections.LineCollection):
    def assemble(self, x, y, color='C0'):
            #assemble line marks as set of segments 
            segments = [np.vstack((vx, vy)).T for vx, vy 
                        in zip(x, y)]
            self.set_segments(segments)
            self.set_color(color)

\end{minted}
In the \mintinline{python}{Line} artist, \mintinline{python}{view} returns a table of edges. Each edge consists of (x,y) points sampled along the line defined by the edge and information such as the color of the edge. As with \mintinline{python}{Point}, the data is then converted into visual variables. In \mintinline{python}{assemble}, this visual representation is composed into a set of line segments, where each segement is the array generated by \mintinline{python}{np.vstack((vx, vy))}. Then the colors of each line segment are set. The colors are guaranteed to correspond to the correct segment because of the atomicity constraint on view.

\begin{figure}[H]
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=\textwidth]{figures/code/bar_v.png}
    \end{subfigure}
    \begin{subfigure}{0.5\textwidth}
        \includegraphics[width=\textwidth]{figures/code/bar_h.png}
    \end{subfigure}
    \caption{Frequency of Penguin types visualized as discrete bars. }
    \label{fig:code_bar_simple}
\end{figure}
The bar charts in figure~\ref{fig:code_bar_simple} are generated with a \mintinline{python}{Bar} artist. The have the same required \vfiber\ components of (position, length), but in \mintinline{pyton}{__init__} of \mintinline{python}{Bar} an additional parameter is set, \texttt{orientation} which controls whether the bars are arranged vertically or horizontally. This paramter only applies holistically to the graphic and never to individual data parameters, and highlights how the model encourages explicit differentiation between parameters in \vtotal\ and graphic parameters applied directly to \vmarkd\.
